Why robot programming is hard

### 2013-07-02 paper.tex
First draft complete

\title {Robot programming for dummies}
\author {Tessa Lau \\
Willow Garage \\
Menlo Park, CA USA \\
{\tt tlau@willowgarage.com}
}
\maketitle

@@

Actually, the problem is not that robots {\em can't} do what we want them to do. The problem is that it's too difficult to {\em tell} robots what we want them to do. The first company to crack the nut of end user robot programming will corner the market on household robots. Human-robot interaction is the missing link for widespread personal robotics.

@@

One potential use case for personal robotics will be in-home care and assistance for disabled adults. The Robots for Humanity project, a collaboration between Willow Garage and Georgia Tech, aims to develop technology that enables persons with motor impairments to perform personal care tasks that they otherwise might not have been able to perform without assistance, such as picking up a dropped object, fetching a towel, or even scratching an itch. Its first prototypes have enabled Henry Evans, a mute quadriplegic, to act in the world for the first time since a brainstem stroke left him paralyzed. Limited to the ability to move a single finger, Henry has been able to instruct Willow Garage's PR2 robots to open cabinets, fetch food from the refrigerator, and even dole out candy to trick-or-treating kids on his behalf.

@@

The Robot Operating System is one of the software advances that supports the rapid development of new robotics technology.  First developed in 2007 at the Stanford Articifial Intelligence Laboratory, ROS was maintained by Willow Garage from 2008-2013 and has recently transitioned into the stewardship of the non-profit Open Source Robotics Foundation.

@@

Concurrent programming is notoriously difficult even for professional programmers due to the possibility for race conditions and the complexity of reasoning about what the program does. Most robot programs can be simplified by only controlling one moving part at a time: first move the base, then raise the torso, then position the left arm, then position the right arm, finally open the cabinet. For many tasks this is a reasonable solution. However there is a subset of human tasks for which we need multiple parts moving simulateously, such as picking up a heavy or bulky object with two hands, or turning one's head to look where one is going.

@@

Another challenge that robot programmers face is specifying how the robot should behave in an uncertain world. The world is constantly changing: any model that a robot could build of where things are located will be obsolete by the time it turns around. People walk around, objects get lost, furniture gets moved. Moreover, limitations in current sensing technology make it difficult to build a perfectly accurate model of the world. For example, differences in lighting make the same object look very different to an RGB camera. Therefore a robot cannot necessarily trust what it perceives, and it must be able to deal with the possibility that what it thinks is true is not actually true.

In addition, robot hardware itself often produces unpredictable results. Unlike software, mechanical parts can slip, burn out, or break. Giving a robot the same instruction multiple times may or may not result in the same result each time.

@@

In order to deal with these various types of uncertainty, professional programmers can learn to use probabilistic models such as Bayesian networks or HMMs, or they write code with lots of built-in error checking. However, casual programmers who just want a robot to perform a job in the home would find it burdensome to have to think through all the ways in which the robot could fail, and specify what the recovery strategy should be in each case.

@@

Why robotics is becoming important
Statistics on size of market
Number of research investments?

Main points:
	EUP is what will enable robotics to spawn new industries
	by enabling businesses to customize robot behavior for their own needs

	Yet programming robots is difficult
	State of the art today is demonstrational, which works well for limited, repetitive tasks
	Or using highly technical tools to specify fine-grained robot behavior
	Behavior specification requires much deep technical knowledge
	Not only the traditionally hard topics in EUP such as conditionals, loops, variables
	But also advanced topics like concurrency, uncertainty, high-dimensionality

The challenge for robotics is to bring robots to market that are consumer-friendly and can operate in human environments. For limited tasks, we are seeing robots that can perform one single task very well without getting in the way of the humans in the household. Vacuum cleaner robots like the iRobot's Roomba and the Neato XV are flooding the market. The Robomow robotic lawn mower performs a similar task for your lawn. Undoubtedly their success is due to their extremely simple interface: the Roomba has a single large button labeled ``Clean''.

read this: http://workshop.iroboticist.com/why-robotics/

\section{Related work}

Roomba, Neato in consumer use
	http://www.neatorobotics.com/
	http://store.irobot.com/home/index.jsp
Baxter human-safe 2-armed manipulator
	runs ROS
	http://www.ros.org/news/2012/09/rethink-ros.html
	http://www.rethinkrobotics.com/index.php/products/baxter/
Industrial robots:
	KUKA robot arm http://en.wikipedia.org/wiki/KUKA
Toyota's human support robot:
	runs ROS
	http://www.gizmag.com/toyota-human-support-robot/24246/
Telepresence robots:
	Beam
	Anybot QB
	Double

\section{Speech interfaces}

Impedance mismatch of speech interfaces

>>>>

\title {Why robot programming is hard}
\author {Tessa Lau \\
Willow Garage \\
Menlo Park, CA USA \\
{\tt tlau@willowgarage.com}
}
\maketitle

@@

Actually, the problem is not that robots {\em cannot} do what we want them to do. The problem is that it is too difficult to {\em tell} robots what we want them to do. The first company to crack the nut of end user robot programming will corner the market on household robots. Human-robot interaction is the missing link for widespread personal robotics.

@@

One potential use case for personal robotics will be in-home care and assistance for disabled adults. The Robots for Humanity project~\cite{rfh}, a collaboration between Willow Garage and Georgia Tech, aims to develop technology that enables persons with motor impairments to perform personal care tasks that they otherwise might not have been able to perform without assistance, such as picking up a dropped object, fetching a towel, or even scratching an itch. Its first prototypes have enabled Henry Evans, a mute quadriplegic, to act in the world for the first time since a brainstem stroke left him paralyzed. Limited to the ability to move a single finger, Henry has been able to instruct Willow Garage's PR2 robots to open cabinets, fetch food from the refrigerator, and even dole out candy to trick-or-treating kids on his behalf.

@@

The Robot Operating System\cite{ros} is one of the software advances that supports the rapid development of new robotics technology.  First developed in 2007 at the Stanford Articifial Intelligence Laboratory, ROS was maintained by Willow Garage from 2008-2013 and has recently transitioned into the stewardship of the non-profit Open Source Robotics Foundation.

@@

Concurrent programming is notoriously difficult even for professional programmers due to the possibility for race conditions and the complexity of reasoning about what the program does. Most robot programs can be simplified by only controlling one moving part at a time: first move the base, then raise the torso, then position the left arm, then position the right arm, finally open the cabinet. For many tasks this is a reasonable solution. However there is a subset of human tasks for which we need multiple parts moving simultaneously, such as picking up a heavy or bulky object with two hands, or turning one's head to look where one is going.

@@

Another challenge that robot programmers face is specifying how a robot should behave in an uncertain world. The world is constantly changing: any model that a robot could build of where things are located will be obsolete by the time it turns around. People walk around, objects get lost, furniture gets moved. Moreover, limitations in current sensing technology make it difficult to build a perfectly accurate model of the world. For example, differences in lighting make the same object look very different to an RGB camera. Therefore a robot cannot necessarily trust what it perceives, and it must be able to deal with the possibility that what it thinks is true is not actually true.

In addition, robot hardware itself often produces unpredictable results. Unlike software, mechanical parts can slip, burn out, or break. Giving a robot the same instruction multiple times may or may not result in the same outcome each time.

@@

In order to deal with these various types of uncertainty, professional programmers can learn to use probabilistic models such as Bayesian networks or HMMs, or they write code with lots of built-in error checking. However, casual programmers who just want a robot to perform a job in the home would find it burdensome to have to elaborate recovery strategies for each of the different failure conditions.

@@

Robot programming environments are beginning to address these challenges and produce useful systems that can be controlled by untrained users. The key tradeoff is simplicity versus expressiveness: in order to make systems that are more tractable for novices, they need to simplify key aspects of the robot's behavior, thereby sacrificing its ability to perform a full range of behavior.

This past generation of commercial home robots illustrates one extreme end of the simplicity/expressiveness tradeoff. iRobot's successful Roomba line of vacuum-cleaning robots\footnote{http://www.irobot.com/us/learn/home/roomba.aspx} sports a single-button interface that simply says ``Clean''.

The next generation of commercial robots in this space will likely look like Rethink Robots' Baxter system\footnote{http://www.rethinkrobotics.com/index.php/products/baxter/}.  Baxter is a two-armed manipulator than can be programmed by demonstration to automate factory work simply by guiding its arms to the appropriate position and tapping icons on the robot's ``face''.  This style of programming makes it possible for people to specify the robot's behavior directly on the robot, without having to deal with any abstractions. In addition, Baxter performs its work in a human-safe way, immediately ceasing motion if it detects that a human has moved into its path, enabling it to work side by side with factory workers unlike previous generations of industrial robots.

For the future, our team at Willow Garage is exploring different tradeoff points in the space of simplicity versus expressiveness. One research direction we have been exploring is to look at specific verticals and design a customized set of robot programming primitives specifically for that domain. Each of these primitives can be hand-crafted in the lab by expert roboticists to embody a rich amount of automation, parallelism, and support for uncertainty. The end user programming problem is then a simpler matter of sequencing these higher-level primitives into useful, customizable programs for their environment.

For example, our team has been doing ethnographic investigations of employees in continuing care retirement centers to discover how robotic helpers could enable them to spend more time with residents. We have discovered that a robotic assistant who could autonomously fetch and deliver supplies to arbitrary locations within the facility could be beneficial to their workflow. Based on those findings, we have defined a set of primitives that include picking up and dropping off supply bins, navigating autonomously to pre-defined locations, and prompting for user input via a touchscreen on the robot's ``face''. End user programmers will be able to sequence these primitives themselves to design custom behaviors for a robotic delivery system, while expert roboticists can focus on making individual primitives robust using the full slate of advanced programming techniques at their disposal.

Over time, as we discover and build new domain-specific primitives, my hope is that we will expand the repertoire of robust end-user-programmable robot behavior. And perhaps that goal of having our own Rosie the Robot in our homes will soon become a reality.

% Why robotics is becoming important
% Statistics on size of market
% Number of research investments?
% 
% Main points:
% 	EUP is what will enable robotics to spawn new industries
% 	by enabling businesses to customize robot behavior for their own needs
% 
% 	Yet programming robots is difficult
% 	State of the art today is demonstrational, which works well for limited, repetitive tasks
% 	Or using highly technical tools to specify fine-grained robot behavior
% 	Behavior specification requires much deep technical knowledge
% 	Not only the traditionally hard topics in EUP such as conditionals, loops, variables
% 	But also advanced topics like concurrency, uncertainty, high-dimensionality
% 
% The challenge for robotics is to bring robots to market that are consumer-friendly and can operate in human environments. For limited tasks, we are seeing robots that can perform one single task very well without getting in the way of the humans in the household. Vacuum cleaner robots like the iRobot's Roomba and the Neato XV are flooding the market. The Robomow robotic lawn mower performs a similar task for your lawn. Undoubtedly their success is due to their extremely simple interface: the Roomba has a single large button labeled ``Clean''.
% 
% 
% read this: http://workshop.iroboticist.com/why-robotics/
% 
% \section{Related work}
% 
% Roomba, Neato in consumer use
% 	http://www.neatorobotics.com/
% 	http://store.irobot.com/home/index.jsp
% Baxter human-safe 2-armed manipulator
% 	runs ROS
% 	http://www.ros.org/news/2012/09/rethink-ros.html
% 	http://www.rethinkrobotics.com/index.php/products/baxter/
% Industrial robots:
% 	KUKA robot arm http://en.wikipedia.org/wiki/KUKA
% Toyota's human support robot:
% 	runs ROS
% 	http://www.gizmag.com/toyota-human-support-robot/24246/
% Telepresence robots:
% 	Beam
% 	Anybot QB
% 	Double
% 
% 
% \section{Speech interfaces}
% 
% Impedance mismatch of speech interfaces
